---
title       : Motor Vehicle Producer Price Index
subtitle    : A Time Series Investigation
author      : Todd Young
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
github:
  user: Young-Todd
  repo: automotive_presentation
---

## Producer Price Indexes 

"A measurement of the average change over time in the selling 
prices received by domestic producers for their output."

-Bureau of Labor Statistics

--- .class #id 

## The Series

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=13, fig.height=7}
library(Quandl)
library(ggplot2)
library(ggfortify)
library(forecast)
library(gridExtra)
library(xts)
library(xtable)
library(MASS)
library(car)
library(tsoutliers)
library(cpm)


Quandl.api_key("Z9Dr3iK6S8q8cuSzomkX")
# Reading in data from Quandl
us_automotive <- Quandl("BLSI/WPU1411")
# Reversing ordering of time (earliest dates first)
us_automotive <- us_automotive[order(us_automotive$Date, -as.numeric(us_automotive$Date)), ]

r_squared <- function(y, y_hat) {
  # Here we are calculating the r-squared
  # value for our models
  
  # Storing the mean of the sample
  y_mean <- mean(y)
  # 
  numerator <- sum( (y-y_hat)^2 )
  denominator <- sum( (y-y_mean)^2 )
  
  r_squared <- 1 - (numerator/denominator)
  return(r_squared)
}

time_index <- seq(from = as.POSIXct("1994-01-01"),to = as.POSIXct("2016-03-01"), by = "month")
prices_xts <- xts(us_automotive$Value, order.by = time_index)

# Creating a time series object 
producer_prices <- ts(us_automotive$Value, start = c(1994,1), end = c(2016,3), frequency = 12)
# Subsetting the series as a 
train <- window(producer_prices, start = 1994, end = 2014)
test <- window(producer_prices, start = 2015)

# Plotting the time series
autoplot(train, main = "Motor Vehicle Producer Price Index")
```

---

## ACF and PACF of the Series

```{r, ACF, fig.keep='none', echo=FALSE}
# ACF plot
acf_plot <- autoplot( acf(train), main = "ACF Plot" )
```

```{r, PACF, fig.keep='none', echo=FALSE}
# PACF plot
pacf_plot <- autoplot( pacf(train), main = "PACF Plot")
```

```{r, ACF & PACF plots, fig.keep='last', fig.height=7, fig.width=13, echo=FALSE}
# Plotting both ACF and PACF plots
grid.arrange(acf_plot, pacf_plot, ncol=2)
```

---

## Series Decomposition

```{r, echo=FALSE, fig.width=13, fig.height=7}
decomposition <- stl(train, "periodic")
# Plotting seasonal decomposition
autoplot(decomposition)
```

---

## A Closer Look at Seasonality

```{r, fig.keep='last', fig.height=7, fig.width=13, echo=FALSE}
us_automotive$month <- factor(strftime(us_automotive$Date,"%b"),levels=month.abb)
# Boxplot of seasonal data
ggplot(us_automotive, aes(x = month, y = Value), main = "Seasonal Boxplots") + geom_boxplot()
```

---

## Holt-Winter's Exponential Smoothing Model

```{r, echo=FALSE, fig.width=13, fig.height=7}
# Creating Holt-Winters' model
holt_winters <- hw(train)
# Creating forecast 15 periods out
holt_winters_forecast <- forecast(holt_winters, h = 15)
# Plotting the smoothing model
autoplot(holt_winters_forecast, main = "Holt-Winters Exponential Smoothing Model")

# Storing accuracy measures for model
accuracy_holt_winters <- as.data.frame(accuracy(holt_winters_forecast, test))
# Saving R-squared for training and test
accuracy_holt_winters$r_squared[1] <- r_squared(train, holt_winters$fitted)
accuracy_holt_winters$r_squared[2] <- NA
colnames(accuracy_holt_winters)[9] <- "R-squared"
```

---

## State Space Model

```{r, echo=FALSE, fig.width=13, fig.height=7}
# Creating state space model
state_space <- ets(train)
# Plotting the state space model
#autoplot(state_space, main = "State Space Model")
# Creating forecast
state_space_forecast <- forecast(state_space, h=13)
# Plotting forecast
autoplot(state_space_forecast, main = "State Space Model Forecast")

# Storing accuracy measures of model
accuracy_state_space <- as.data.frame( accuracy(state_space_forecast, test) )
# Caculating R-squared for training and test
accuracy_state_space$r_squared[1] <- r_squared(train, state_space$fitted)
accuracy_state_space$r_squared[2] <- r_squared(test, state_space_forecast$mean)
colnames(accuracy_state_space)[9] <- "R-squared"
```

---

## Seasonal Arima Model

```{r, echo=FALSE, fig.width=13, fig.height=7}
# Fitting a seasonal Arima model
arima_model <- auto.arima(train) 
# Plotting the model
#autoplot(arima_model, main = "Seasonal Arima Model")
# Creating forecast
arima_forecast <- forecast(arima_model, h=15)
# Plotting forecast
autoplot(arima_forecast, main = "Seasonal ARIMA Model Forecast")

# Storing accuracy measures
accuracy_arima <- as.data.frame( accuracy(arima_forecast, test) )
# Finding R-squared for training and test
accuracy_arima$r_squared[1] <- r_squared(train, arima_model$x)
accuracy_arima$r_squared[2] <- r_squared(test, arima_forecast$mean)
colnames(accuracy_arima)[9] <- "R-squared"
```

--- 

## Model Evaluations

```{r, results='asis', message=FALSE, fig.width=13, fig.height=7, echo=FALSE}
print( xtable(accuracy_holt_winters, caption = "Holt-Winter's Model Accuracy Measures"), type = 'html')
```

```{r, results='asis', message=FALSE, fig.width=13, fig.height=7, echo=FALSE}
print( xtable(accuracy_state_space, caption = "State Space Model Accuracy Measures"), type = 'html')
```

```{r, results='asis', message=FALSE, echo=FALSE}
print( xtable(accuracy_arima, caption = "Seasonal Arima Model Accuracy Measures"), type = 'html')
```

---